---
title: "N-1 Interactions by Trial Type and Reaction Time"
description: |
  Four way interactions between my pre-trial variables I am interested in, n-1 CSI and n-1 Congruency by currect reaction time and trial type.     
author:
  - name: Sophia Angleton 
    url: https://doi.org/10.17605/OSF.IO/4DVGS
    affiliation: University of Oregon
    affiliation_url: https://www.uoregon.edu/
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


```

```{r}
#| echo: False
xaringanExtra::use_panelset()
```

```{r}
#| include: False
library(here)
library(tidyverse)
library(ez)
library(psych)
library(sparkline)
library(readr)
library(ggsignif)
library(car)
library(broom)
library(sjPlot)

PrepSwitch <- read.csv("~/EDLD_652/FInal Blog/_posts/2025-02-19-first-post/PrepSwitch.csv")
theme_set(theme_minimal())
head(PrepSwitch)

#clean data 

#removing practice trials from our df
prep <- PrepSwitch %>% 
  filter(practice != 1)

sum_er <- prep %>% #determine 80% accuracy 
  group_by(id) %>% 
summarize(avg_er = mean(error)) %>% 
  print()

sum_er <- sum_er %>% 
  mutate(outlier_er =  (avg_er > 0.20)) %>% 
  print() #2 people fall below 80% accuracy. id 19 and 30. 

#removing those outliers here and adding switch trials and lag trials and incongruencies. 

prep <- prep %>% 
  filter(id != 19, id != 30) %>% 
  mutate(switch = ifelse(task == lag(task,1), "No-switch", "Switch"), 
         lcsi = lag(csi, 1), incongruent = ifelse(numpresent %in% c(1,3,6,8), 0, 1), lincongruent = lag(incongruent,1), lerror = lag(error, 1), lrt = lag(rt, 1)) %>%
  mutate(lcsi = ifelse(lcsi == 1, "Short", "Long"), switch = as.factor(switch), lincongruent = ifelse(lincongruent == 0, "Congruent", "Incongruent")) %>% #lag looks at trial before, need to filter out the first of each block bc nothing to compare
  filter(trial != 1, rt > 200 & rt < 4000)

prep <- prep %>% 
  group_by(id) %>% 
  mutate(alltrials = n()) %>% 
  ungroup() #I noticed there was one subject that only had 229 trials total...

#view(prep)

```

::::: {.panelset}

::: {.panel}

## First Visualization {.panel-name}

I want to look at the interaction effect between n-1 CSI level, n-1 congruence, and current reaction time by trial type.  

```{r}

preplm = lm(rt ~ lcsi*switch*lincongruent, data = prep)

a<- plot_model(preplm, type = "pred", 
           terms = c("lincongruent", # x-axis
                     "switch" ,# color of lines
                     "lcsi"))
a
```

:::

::: {.panel}

## Second Visualization {.panel-name}

I want to change around where my variables go and add geom_smooth(). 

```{r}

preplm = lm(rt ~ lcsi*switch*lincongruent, data = prep)

b<- plot_model(preplm, type = "pred", 
           terms = c("lcsi", # x-axis
                     "lincongruent" ,# color of lines
                     "switch")) +
  geom_smooth(method = "lm")
b

```

:::

::: {.panel}

## Third Visualization {.panel-name}

I want to add in some custom colors for my color of lines but still keep it grouped by congruence.

```{r}
custom_colors <- c( "Congruent" = "#009EB0","Incongruent" = '#C44601') #color blind safe pallet

 preplm = lm(rt ~ lcsi*switch*lincongruent, data = prep)

c<- plot_model(preplm, type = "pred", 
           terms = c("lcsi", # x-axis
                     "lincongruent" ,# color of lines
                     "switch")) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = custom_colors)
c
```

:::

::: {.panel}

## Final Visualization {.panel-name}

Finally, I want to make my titles and axis neat and adjust the y-axis coordinates. 

```{r}
custom_colors <- c( "Congruent" = "#009EB0","Incongruent" = '#C44601') #color blind safe pallet

preplm = lm(rt ~ lcsi*switch*lincongruent, data = prep)

d<- plot_model(preplm, type = "pred", 
           terms = c("lcsi", # x-axis
                     "lincongruent" ,# color of lines
                     "switch")) +
  geom_smooth(method = "lm") +
    coord_cartesian(ylim = c(700, 900))+
  scale_color_manual(values = custom_colors) +
    labs(title = "Predicted Values of Response Time",
       x = "n-1 CSI",
       y = "Response Time (msec)",
       color = "n-1 Congruence"
       ) 
d
```

:::

:::::
