---
title: "first-post"
description: |
  A short description of the post.
author:
  - name: Sophia Angleton
    url: https://example.com/norajones
date: 2025-02-19
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

I will use the "EDFacts_math_achievement_lea_2010_2019" to make my data visualizations on a dashboard format. I am interested in this dataset becuase it centralizes data provided by LEA or the local education agency. I want to focus on the academic acheivement in mathmatics becuase I feel like I can get a better understanding of performance trends by looking at just one educational area.



```{r}
#| include: False
library(tidyverse)
library(edld652)
library(janitor)

#set_key(1234, overwrite = T)

get_documentation("EDFacts_math_achievement_lea_2010_2019")

edu_math <- get_data("EDFacts_math_achievement_lea_2010_2019")

edu_math <- clean_names(edu_math)

theme_set(theme_minimal())
```

```{r}
plot1 <- edu_math %>% 
  filter(!is.na(ecd_mth03pctprof)) %>% 
  ggplot(aes(ecd_mth03numvalid, ecd_mth03pctprof)) +
  geom_col() +
  labs(title = 'Academic Acheivement by Financial Status', subtitle = 'Mathmatic Acheivement at The Third-grade level', x = 'Economically Disadvantaged Students', y = "Percent scoring at or above proficiency level")
plot1

plot2 <- edu_math %>% 
  filter(!is.na(ecd_mth08pctprof)) %>% 
  ggplot(aes(ecd_mth08numvalid, ecd_mth08pctprof)) +
  geom_col() +
  labs(title = 'Academic Acheivement by Financial Status', subtitle = 'Mathmatic Acheivement at The Eighth-grade level', x = 'Economically Disadvantaged Students', y = "Percent scoring at or above proficiency level")
plot2
```

