---
title: "Trial Type, Response Time and Error Rate"
description: |
 We are interested in trial type, switch or no-switch, and its effect on reaction time and error rate.
author:
  - name: Sophia Angleton
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

I want to look at the interaction effects between switch trials, or when someone is switching from one task instruction to another, and no-switch trials, or when someone is doing one task repeatedly. I am doing this by looking at switch costs, represented through differences in response time.

+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **What the variables mean:**                                                                                                                                                 |
+==============================================================================================================================================================================+
| -   Task: 1 is parity (odd or even) , 2 is size (bigger or smaller than 5)                                                                                                   |
|                                                                                                                                                                              |
| -   rt: response time                                                                                                                                                        |
|                                                                                                                                                                              |
| -   Error: 0 = no error, 1 = yes error                                                                                                                                       |
|                                                                                                                                                                              |
| -   Trial type: 1 = switch trials, or when someone is switching from one task instruction to another, and 0 = no-switch trials, or when someone is doing one task repeatedly |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r}
#| echo: False
xaringanExtra::use_panelset()
```

```{r}
#| include: False
library(here)
library(tidyverse)
library(ez)
library(psych)
library(sparkline)
library(readr)
library(ggsignif)
library(car)
library(broom)

PrepSwitch <- read.csv("~/EDLD_652/FInal Blog/_posts/2025-02-19-first-post/PrepSwitch.csv")
theme_set(theme_minimal())
head(PrepSwitch)

#clean data 

#removing practice trials from our df
prep <- PrepSwitch %>% 
  filter(practice != 1)

sum_er <- prep %>% #determine 80% accuracy 
  group_by(id) %>% 
summarize(avg_er = mean(error)) %>% 
  print()

sum_er <- sum_er %>% 
  mutate(outlier_er =  (avg_er > 0.20)) %>% 
  print() #2 people fall below 80% accuracy. id 19 and 30. 

#removing those outliers here and adding switch trials and lag trials and incongruencies. 

prep <- prep %>% 
  filter(id != 19, id != 30) %>% 
  mutate(switch = ifelse(task == lag(task,1), "No-switch", "Switch"), 
         lcsi = lag(csi, 1), incongruent = ifelse(numpresent %in% c(1,3,6,8), 0, 1), lincongruent = lag(incongruent,1), lerror = lag(error, 1), lrt = lag(rt, 1)) %>%#lag looks at trial before, need to filter out the first of each block bc nothing to compare
  filter(trial != 1, rt > 200 & rt < 4000)

prep <- prep %>% 
  group_by(id) %>% 
  mutate(alltrials = n()) %>% 
  ungroup() #I noticed there was one subject that only had 229 trials total...

#view(prep)


#aggragation of the parts of our data that we are interesting in seeing

agglag <- prep %>% 
  group_by(id, lcsi, switch) %>% #agg lag groups to then do anova. 
  summarize(rt = mean(rt[error == 0], na.rm = T), 
            error = mean(error, na.rm = T)) 
agglag

agglag2 <- agglag %>% #this is the important thing we want to see in the n-1 trials. looking at pattern of means here 
  group_by(lcsi, switch) %>% 
  summarize(rt = mean(rt, na.rm = T), 
            error = mean(error, na.rm = T)) 
agglag2


#doing our anovas for the agglag aggragates: 

anovart <- ezANOVA(data = agglag ,  
dv = rt, wid = id, within = .(lcsi, switch), type = 3, detailed = T)
anovart

anovaer <- ezANOVA(data = agglag, 
dv = error, wid = id, within = .(lcsi, switch), type = 3, detailed = T)
anovaer
```

:::::: panelset
::: panel
## First Visualization {.pannel-name}

```{r}

boxplot <- agglag %>% 
  ggplot(aes(x = switch, y = rt)) +
  geom_boxplot()
boxplot

```
:::

::: panel
## By Reaction Time {.pannel-name}

```{r}

boxplot2 <- agglag %>% 
  ggplot(aes(x = switch, y = rt)) +
  geom_boxplot(fill = "#AABA99") +
  labs(title = "Trial Type On Reaction Time", x = "Trial Type", y= "Response Time (msec)", subtitle = "Mean Response Times by Trial Type")
boxplot2

```
:::

::: panel
## By Error {.panel-name}

Lets look at this by error.

```{r}
boxploter <- agglag %>% 
  ggplot(aes(x = switch, y = error)) +
  geom_boxplot(fill = "#AABA99") +
  labs(title = "Trial Type On Error Rate", x= "Trial Type", y= "Error Rate", subtitle = "Mean Error Rates by Trial Type")
boxploter
```
:::
::::::

## Interpretation

::::: panelset
::: panel
## Trial Type on Reaction Time {.panel-name}

We see that there is a difference in reaction time by trial types, where people tend to have a shorter reaction time in no-switch trials compared to switch trials. However, we see that there are two outliers, meaning that there are two subjects who had an unusually high mean response time in switch trials than others. This may contribute to skewing our switch trial mean response time to be longer than it should be. This is what we expect to see according to previous understanding of trial type and response time.
:::

::: panel
## Trial Type on Error Rate {.panel-name}

We see that error rate does change from no-switch to switch trials, where no-switch trials seem to lead to lower average error rates by subject, and vice versa for switch trials. We see an outlier in the no-switch trial where one subject had unusually high error rates when completing no-switch trials which may be skewing the plot higher in overall error rate for the no-switch trial type. This is what we expect to see according to previous undersanding of trial type and error rate.
:::
:::::
