---
title: "Trial Type and Error Rate"
description: |
  We are interested in trial type, switch or no-switch, and its effect on error rate. 
author:
  - name: Sophia Angleton
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Trial Type and Error Rate

Secondly, I want to look at the interaction effects between trial type and the error rate. I want to do this to account for a possible speed-accuracy tradeoff, where as you get faster in a task you may be less accurate. This is important to account for because it is not the tradeoff we are interested in testing and want to check that this is not the case. 

```{r}
#| echo: False
xaringanExtra::use_panelset()
```

```{r}
#| include: False
library(here)
library(tidyverse)
library(ez)
library(psych)
library(sparkline)
library(readr)
library(ggsignif)
library(car)
library(broom)

PrepSwitch <- read.csv("~/EDLD_652/FInal Blog/_posts/2025-02-19-first-post/PrepSwitch.csv")
theme_set(theme_minimal())
head(PrepSwitch)

#clean data 

#removing practice trials from our df
prep <- PrepSwitch %>% 
  filter(practice != 1)

sum_er <- prep %>% #determine 80% accuracy 
  group_by(id) %>% 
summarize(avg_er = mean(error)) %>% 
  print()

sum_er <- sum_er %>% 
  mutate(outlier_er =  (avg_er > 0.20)) %>% 
  print() #2 people fall below 80% accuracy. id 19 and 30. 

#removing those outliers here and adding switch trials and lag trials and incongruencies. 

prep <- prep %>% 
  filter(id != 19, id != 30) %>% 
  mutate(switch = ifelse(task == lag(task,1), "No-switch", "Switch"), 
         lcsi = lag(csi, 1), incongruent = ifelse(numpresent %in% c(1,3,6,8), 0, 1), lincongruent = lag(incongruent,1), lerror = lag(error, 1), lrt = lag(rt, 1)) %>%#lag looks at trial before, need to filter out the first of each block bc nothing to compare
  filter(trial != 1, rt > 200 & rt < 4000)

prep <- prep %>% 
  group_by(id) %>% 
  mutate(alltrials = n()) %>% 
  ungroup() #I noticed there was one subject that only had 229 trials total...

#view(prep)


#aggragation of the parts of our data that we are interesting in seeing

agglag <- prep %>% 
  group_by(id, lcsi, switch) %>% #agg lag groups to then do anova. 
  summarize(rt = mean(rt[error == 0], na.rm = T), 
            error = mean(error, na.rm = T)) 
agglag

agglag2 <- agglag %>% #this is the important thing we want to see in the n-1 trials. looking at pattern of means here 
  group_by(lcsi, switch) %>% 
  summarize(rt = mean(rt, na.rm = T), 
            error = mean(error, na.rm = T)) 
agglag2


#doing our anovas for the agglag aggragates: 

anovart <- ezANOVA(data = agglag ,  
dv = rt, wid = id, within = .(lcsi, switch), type = 3, detailed = T)
anovart

anovaer <- ezANOVA(data = agglag, 
dv = error, wid = id, within = .(lcsi, switch), type = 3, detailed = T)
anovaer
```


::::: {.panelset}

::: {.panel}

## First Visualization {.pannel-name}

```{r}
boxploter <- agglag %>% 
  ggplot(aes(x = switch, y = error)) +
  geom_boxplot() 
boxploter
```
:::

::: {.panel}

## Final Visualization {.pannel-name} 

```{r}
boxploter <- agglag %>% 
  ggplot(aes(x = switch, y = error)) +
  geom_boxplot(fill = "#AABA99") +
  labs(title = "Trial Type On Error Rate", x = "Trial Type", y= "Error", subtitle = "Mean Error Rates by Trial Type")
boxploter
```

:::

::: {.panel}

## Interpretation {.panel-name}

We see that error rate does change from no-switch to switch trials, where no-switch trials seem to lead to lower average error rates by subject, and vice versa for switch trials. We see an outlier in the no-switch trial where one subject had unusually high error rates when completing no-switch trials which may be skewing the plot higher in overall error rate for the no-switch trial type. This is what we expect to see according to previous undersanding of trial type and error rate.

:::

:::::
