---
title: "Trial Type, Cue Stimulus Interval, and Response Time"
description: |
  I want to use an interaction between two catagorical variables, trial type and cue stimulus interval, and plot them by response time. 
author:
  - name: Sophia Angleton
output:
  distill::distill_article:
    self_contained: false
---

+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| What the variables mean:                                                                                                                                                     |
+==============================================================================================================================================================================+
| -   CSI or cue-stimulus interval: 1 is short (no preparation), 2 is long (yes preparation)                                                                                   |
|                                                                                                                                                                              |
| -   Task: 1 is parity (odd or even) , 2 is size (bigger or smaller than 5)                                                                                                   |
|                                                                                                                                                                              |
| -   rt: response time                                                                                                                                                        |
|                                                                                                                                                                              |
| -   Trial type: 1 = switch trials, or when someone is switching from one task instruction to another, and 0 = no-switch trials, or when someone is doing one task repeatedly |
+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
#| echo: False
xaringanExtra::use_panelset()
```

```{r}
#| include: False
library(here)
library(tidyverse)
library(ez)
library(psych)
library(sparkline)
library(readr)
library(ggsignif)
library(car)
library(broom)

PrepSwitch <- read.csv("~/EDLD_652/FInal Blog/_posts/2025-02-19-first-post/PrepSwitch.csv")
theme_set(theme_minimal())
head(PrepSwitch)

#clean data 

#removing practice trials from our df
prep <- PrepSwitch %>% 
  filter(practice != 1)

sum_er <- prep %>% #determine 80% accuracy 
  group_by(id) %>% 
summarize(avg_er = mean(error)) %>% 
  print()

sum_er <- sum_er %>% 
  mutate(outlier_er =  (avg_er > 0.20)) %>% 
  print() #2 people fall below 80% accuracy. id 19 and 30. 

#removing those outliers here and adding switch trials and lag trials and incongruencies. 

prep <- prep %>% 
  filter(id != 19, id != 30) %>% 
  mutate(switch = ifelse(task == lag(task,1), "No-switch", "Switch"), 
         lcsi = lag(csi, 1), incongruent = ifelse(numpresent %in% c(1,3,6,8), 0, 1), lincongruent = lag(incongruent,1), lerror = lag(error, 1), lrt = lag(rt, 1)) %>% #lag looks at trial before, need to filter out the first of each block bc nothing to compare
  filter(trial != 1, rt > 200 & rt < 4000)

prep <- prep %>% 
  group_by(id) %>% 
  mutate(alltrials = n()) %>% 
  ungroup() #I noticed there was one subject that only had 229 trials total...

view(prep)


#aggragation of the parts of our data that we are interesting in seeing

agglag <- prep %>% 
  group_by(id, switch, csi) %>% #agg lag groups to then do anova. 
  summarize(rt = mean(rt[error == 0], na.rm = T), 
            error = mean(error, na.rm = T)) 
agglag

```

:::::: panelset
::: panel
## First Visualization {.pannel-name}

```{r}
# Create an interaction between the two factors
agglag$switchcsi <- interaction(agglag$switch, agglag$csi)
 
custom_colors <- c("#009EB0", "#AABA99", '#C44601', '#F194B8') #color blind safe pallet
 

plot <- ggplot(agglag, aes(x = switchcsi, y = rt, fill = switchcsi)) +
  geom_violin() +
  scale_fill_manual(values = custom_colors)
plot

```
:::

::: panel
## Second Visualization {.pannel-name}

I want to add an overlay of my individual data points by subject on my violin plots. I also want to remove the redundant legend.

```{r}
 
plot2 <- ggplot(agglag, aes(x = switchcsi, y = rt, fill = switchcsi)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, size = 3, alpha = 0.8) +
  scale_fill_manual(values = custom_colors) +
  theme(
    legend.position = "none"
  )
plot2
```
:::

::: panel
## Final Visualization {.panel-name}

Adding in some better visualization the plot colors, sizing, and text.

```{r}

custom_labels <- c("No-switch.1" = "NS & Short CSI", "Switch.1" = "S & Short CSI", "No-switch.2" = "NS & Long CSI", "Switch.2" = "S & Long CSI")

finalplot <- ggplot(agglag, aes(x = switchcsi, y = rt, fill = switchcsi)) +
  geom_violin(width = 0.5, alpha = 0.6, color = "#FFFDD0") +
  geom_jitter(width = 0.2, height = 0, size = 1.7, alpha = 0.8, color = '#363737') +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Trial Type and Cue Stimulus Interval (CSI) by Response Time", subtitle = "NS = No-Switch, S = Switch", x = "", y = "Response Time (msec)") +
    scale_x_discrete(labels = custom_labels) +
  theme(legend.position = "none")
finalplot
```
:::
::::::

## Interpretation

Visually, it doesn't seem like there is much of a change in average reaction time when grouped by trial type (no-switch or switch trial) and cue stimulus interval (short or long preparation before onset of stimulus). Indeed, it seems that through an anova of all these aggregated variables, we see that there is not a significant differences in the pattern of reaction time means in any of these interactions. The biggest difference we do see visually is switch trials and a short CSI- where supposedly when subjects are the most flexible, their reaction times skew longer than in other interactions.
