---
title: "Interaction Plots"
description: |
  I am looking at the interactions between trial type, switch or no-switch, and two main effect variables, the cue stimulus interval (CSI) of the trial or the congruency of the trial. 
author:
  - name: Sophia Angleton
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
#| echo: False
xaringanExtra::use_panelset()
```

```{r}
#| include: False
library(here)
library(tidyverse)
library(ez)
library(psych)
library(readr)
library(ggsignif)
library(plotly)
library(ggthemes)
library(patchwork)
PrepSwitch <- read.csv("~/EDLD_652/FInal Blog/_posts/2025-02-19-first-post/PrepSwitch.csv")

#clean data 
theme_set(theme_solarized())
#removing practice trials from our df
prep <- PrepSwitch %>% 
  filter(practice != 1)

sum_er <- prep %>% #determine 80% accuracy 
  group_by(id) %>% 
summarize(avg_er = mean(error)) %>% 
  print()

sum_er <- sum_er %>% 
  mutate(outlier_er =  (avg_er > 0.20)) %>% 
  print() #2 people fall below 80% accuracy. id 19 and 30. 

#removing those outliers here and adding switch trials and lag trials and incongruencies. 

prep <- prep %>% 
  filter(id != 19, id != 30) %>% 
  mutate(switch = ifelse(task == lag(task,1), "No-Switch", "Switch"), 
         lcsi = lag(csi, 1), incongruent = ifelse(numpresent %in% c(1,3,6,8), 0, 1), lincongruent = lag(incongruent,1), lerror = lag(error, 1), lrt = lag(rt, 1)) %>%#lag looks at trial before, need to filter out the first of each block bc nothing to compare
  filter(trial != 1, rt > 200 & rt < 4000)

prep <- prep %>% 
  group_by(id) %>% 
  mutate(alltrials = n()) %>% 
  ungroup() #I noticed there was one subject that only had 229 trials total...

view(prep)

```

::::::: panelset
::: panel
## First Visualization {.pannel-name}

First interaction between trial type and congruence. 

```{r}
# First interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$incongruent,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#AABA99"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o") # Custom y-axis label
```
:::

::: panel
## Second Visualization {.pannel-name}

Second interaction between trial type and cue stimulus interval (CSI). 

```{r}
# Second interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$csi,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#198BA9"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o" 
                 )                  # Blank y-axis label
```
:::

::: panel
## Third Visualization {.pannel-name}

I want to add these two plots together.

```{r}
# Set up the plotting area with two plots side by side
par(mfrow = c(1, 2))

# First interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$incongruent,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#AABA99"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o") 


# Second interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$csi,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#198BA9"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o")                   


# Reset plotting parameters to default
par(mfrow = c(1, 1))
```
:::

::: panel
## Final Visualization {.pannel-name}

I want to customize the titles and axis.

```{r}
# Set up the plotting area with two plots side by side
par(mfrow = c(1, 2))

# First interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$incongruent,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#AABA99"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o",               # No border for legend
                 xlab = "",         # Custom x-axis label
                 ylab = "Response Time (msec)", 
                 legend = F) # Custom y-axis label

# Add custom legend to the first plot
legend("topleft", legend = c("Congruent", "Incongruent"), col = c("black", "#AABA99"), pch = c(19, 17), bty = "o", cex = .7)
# Manually add the trace label centered above the first plot
mtext("Congruency", side = 3, line = .3, at = mean(par("usr")[1:2]), cex = .9)
# Second interaction plot without the automatic trace label and legend
interaction.plot(x.factor = prep$switch,
                 trace.factor = prep$csi,
                 response = prep$rt,
                 fun = mean,
                 type = "b",
                 col = c("black", "#198BA9"),
                 pch = c(19, 17),
                 fixed = TRUE,
                 trace.label = "",          # Remove trace factor label
                 leg.bty = "o",               # No border for legend
                 xlab = "",                   # Blank x-axis label
                 ylab = "", 
                 legend = F
                 )                   # Blank y-axis label

# Add custom legend to the second plot
legend("topleft", legend = c("Short CSI", "Long CSI"), col = c("black", "#198BA9"), pch = c(19, 17), bty = "o", cex = .7)
mtext("Cue Stimulus Interval", side = 3, line = .3, at = mean(par("usr")[1:2]), cex = .9)

# Manually add the main title centered above both plots
mtext("Interactions Of Congruency and CSI On Trial Type", side = 3, line = 2, at = .4, cex = 1.5)
mtext("Trial Type", side = 1, line = 3, at = .4, cex = 1.1)

# Reset plotting parameters to default
par(mfrow = c(1, 1))
```
:::
:::::::

## Interpretation

### *Congruence*

There seems to be negligible differences between congruence/incongruity when going from a no-switch trial to a switch trial. However, when looking at the trends for each respective effect, we see things that make sense. Congruent trials, or trials were the task instruction (size or parity) does not matter, should be lower in average response time than incongruity which is what we see. Furthermore, we see that no-switch congruence leads to the shortest response times and switch incongruity (size and task instruction does matter) leads to the longest response times.

### *Cue Stimulus Interval*

Similarly to the congruence interaction plot, we see that between the short and long CSI's there is little to no interaction effect. HOWEVER we find a suprising result: We see when CSI is long (meaning long preparation), response times are lower on average in both no-switch and switch conditions. This is suprising because longer preparation would induce more cognitive stability (theoretically) and thus lead to a longer switch cost (seen in reaction time) than short CSI. This may be some evidence toward a anti-tradeoff model being represented. 